---
title: "Brno Crime"
subtitle: "[DATASET]"
date: "`r format(Sys.Date())`"

weight: 3

summary: Visualization of crime in Brno.

output:
  md_document:
    preserve_yaml: TRUE

image:
  # Focal point (optional)
  # Options: Smart, Center, TopLeft, Top, TopRight, Left, Right, BottomLeft, Bottom, BottomRight
  focal_point: "Smart"

tags:
  - "brno"
  - "dataset"
  - "open-data"
  - "crime"
---

Let's start by including a few libraries ...


```{r knitr-setup, include=FALSE}
library(tidyverse)
library(rmarkdown)

PROJECT_NAME <- metadata$title %>% tolower() %>% str_replace_all(pattern = ' ', replacement = '-')

knitr::opts_knit$set(base.dir = paste0(getwd(), '/project/', PROJECT_NAME))
knitr::opts_chunk$set(echo = T, collapse = T,
                      fig.path = paste0("project/", PROJECT_NAME, "/figure/"),
                      fig.width = 11, fig.height = 7,
                      results = 'hide', message = FALSE, warning = FALSE)
```

```{r setup}
library(tidyverse)
library(magrittr)
library(wrapr)

library(cowplot)
library(ggplot2)

library(plotly)
library(formattable)

# custom color scale
# inspired by https://github.com/drsimonj/blogR/blob/master/Rmd/creating_corporate_colors_ggplot2.Rmd
source(file = "code/drsimonj.R", local = T)

# set default ggplot theme due to cowplot usage
theme_set(theme_grey())
```

Define some helper functions for future uusage

```{r}
get_colname <- function(.data, idx) { colnames(.data)[idx] %>% as.name() }
get_density <- function(x, ...) {
  dens <- density(x, ...)
  ix <- findInterval(x, dens$x)
  return(dens$y[ix])
}

# ---- Types of crimes in our data ----

assign_crime_cat <- function(v, ...) {
  sapply(v, function(x) {
    case_when(
      x < 200 ~ "Violent crime",
      x < 300 ~ "Sexual harassment",
      x < 500 ~ "Robbery & Burglary",
      x < 580 ~ "Fraud",
      x < 630 ~ "Vandalism",
      x < 690 ~ "Drugs, Weapons & Dealing",
      x < 800 ~ "Other criminal activities",
      x < 900 ~ "Financial and economical criminal acts",
      x >= 900 ~ "Military crimes"
    )
  })
}
```

Get the data for the crime

> Data is available at: [DATA.BRNO](https://data.brno.cz/dataset/?id=kriminalita-v-brne)

```{r}
dat <- readxl::read_xlsx("data/criminality/raw/trestni_cinnost_2016.xlsx")
```

This data is not very polished

```{r, echo=FALSE, results='asis'}
as.data.frame(head(dat)) %>%
  knitr::kable(caption = "Brno crime data")
```

Let's fix that a little bit

```{r}
# fix names
colnames(dat) %<>% make.names()

# filter rows where all columns are NA
dat %<>%
  filter_all(any_vars(!is.na(.))) %>% as.data.frame() %>%
  transform(`Celková.způsobená.škoda.v.Kč` = currency(
    `Celková.způsobená.škoda.v.Kč`, symbol = "CZK", digits = 0L, sep = " "))

col_e <- readxl::read_xls("data/criminality/raw/sloupec_e.xls")
col_g <- readxl::read_xls("data/criminality/raw/sloupec_g.xls")
col_h <- readxl::read_xls("data/criminality/raw/sloupec_h.xls")
col_i <- readxl::read_xls("data/criminality/raw/sloupec_i.xls")
```


Categorize crimes

```{r}
dat %<>%
  group_by(Takticko.statistická.klasifikace.činu) %>%
  mutate(crime.category = factor(assign_crime_cat(Takticko.statistická.klasifikace.činu)))

## reorder factors
dat$crime.category %<>% factor(unique(assign_crime_cat(seq(1, 1000, 50))))
```

Plot crimes by category

```{r}
plot.crimes.by.category <- ggplot(dat, aes(x = factor(Takticko.statistická.klasifikace.činu))) +
  geom_bar(aes(fill = crime.category)) +
  theme(
    axis.line.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),

    panel.background = element_blank(),

    plot.title = element_text(hjust = 0.5)
  ) +
  ylab("Number of crimes") +
  labs(
    fill = "Crime category",
    title = "Crime in Brno by Category"
  )
```

```{r, echo=FALSE}
plot.crimes.by.category
```


## Damage caused by crime incidents

```{r}
dat.damage <- dat %>%
  select(Základní.útvar.PČR..kde.došlo.k.činu,
         Celková.způsobená.škoda.v.Kč,
         crime.category) %>%
  filter(str_detect(crime.category, "Robbery|Fraud"))
```

Initial exploration

```{r}
ggplot(dat.damage, aes(x = '', y = Celková.způsobená.škoda.v.Kč)) +
  geom_jitter(width = 0.2) +  # add some jitter to see the underlying data more clearly
  scale_y_continuous(labels = function(x, ...) {
    currency(x, ..., digits = 0L, symbol = "CZK", sep = ' ')
  }) +
  ylab("Total damage in CZK") +
  labs(title = "Damage caused by crime incidents") +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),

    plot.title = element_text(hjust = 0.5),

    panel.background = element_blank()
  )
```

As we can see, there are some incredible outliers, let's get rid of them by selecting .95p

```{r}
# 95% quantile
dat.95q <- dat[which(dat$Celková.způsobená.škoda.v.Kč < quantile(dat$Celková.způsobená.škoda.v.Kč, .95)),]

# density of the data points
dens.95q <- get_density(dat.95q$Celková.způsobená.škoda.v.Kč, adjust = 5, n = 50)
```

And give it another shot, this time with coloring by density.

```{r}
plot.scatter <- ggplot(dat.95q, aes(x = `Celková.způsobená.škoda.v.Kč`, y = '')) +
  geom_jitter(aes(color = dens.95q), width = 0.2, show.legend = F) +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),

    axis.title.x = element_text(
      margin = margin(t = 10)),

    panel.background = element_blank(),

    plot.title = element_text(
      hjust = 0.5, vjust = 0.2, margin = margin(b = 20))
  ) +
  scale_x_continuous(
    breaks = {
      dat.95q %>% select(Celková.způsobená.škoda.v.Kč) %.>% seq(min(.), max(.), by=25000) %>% round(1)
    },
    labels = function(x, ...) {
      currency(x, ..., digits = 0L, symbol = "CZK", sep = ' ')
    }) +
  xlab("Damage in CZK") +
  labs(title = "Damage caused by crime incidents")
```

```{r, echo=TRUE}
plot.scatter
```

## Marginal density plot

```{r}
plot.density <- ggplot(dat.95q, aes(x = `Celková.způsobená.škoda.v.Kč`)) +
  geom_density() +
  theme(
    axis.line.y = element_blank(),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),

    panel.background = element_blank(),

    plot.title = element_text(
      hjust = 0.5, vjust = 0.2, margin = margin(b = 20))
  )
```

```{r, echo=TRUE}
ggdraw(insert_xaxis_grob(plot.scatter, plot.density))
```


## Comparison between Facets

```{r}
dat.95q <- dat.95q %>% group_by(Základní.útvar.PČR..kde.došlo.k.činu) %>%
  mutate(mu = mean(Celková.způsobená.škoda.v.Kč)) %>%
  mutate(med = median(Celková.způsobená.škoda.v.Kč))

# aquire total damage done by the place of crime
dat.total <- dat.95q %>%
  group_by(Základní.útvar.PČR..kde.došlo.k.činu) %>%
  mutate(total.damage = sum(Celková.způsobená.škoda.v.Kč, na.rm = T) / 1e6) %>%
  mutate(total.damage.ratio = sum(Celková.způsobená.škoda.v.Kč, na.rm = T) / n()) %>%
  mutate(total.incidents = n()) %>%
  select(
    Základní.útvar.PČR..kde.došlo.k.činu,
    total.damage, total.damage.ratio, total.incidents
  ) %>%
  unique()

```

```{r}
plot.jitter <- ggplot(dat.95q, aes(x = '', y = `Celková.způsobená.škoda.v.Kč`)) +
  geom_jitter(aes(color = dens.95q), width = 0.2, show.legend = F) +
  facet_wrap(vars(`Základní.útvar.PČR..kde.došlo.k.činu`), strip.position = 'bottom') +
  geom_hline(  # total mean line
    aes(yintercept = mu),
    alpha = 0.4,
    linetype = 'dashed') +
  ylab("Damage in CZK") +
  labs(
    title = "Damage caused by crime incidents") +
  theme(
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(
      margin = margin(r = 10)
    ),
    plot.title = element_text(
      hjust = 0.5, vjust = 0.2, margin = margin(b = 20))
  )
```


```{r, echo=FALSE}
plot.jitter
```

```{r}
plot.dtotal <- ggplot(dat.total, aes(x = Základní.útvar.PČR..kde.došlo.k.činu,
                      y = total.damage)) +
  # geom_col(aes(fill = Základní.útvar.PČR..kde.došlo.k.činu)) +
  geom_col(aes(fill = total.incidents)) +
  geom_hline(  # total mean line
    aes(yintercept = mean(total.damage)),
    alpha = 0.4,
    linetype = 'dashed') +
  scale_fill_drsimonj(palette = 'cool', discrete = F) +
  ylab("Damage in millions (CZK)") +
  xlab("Brno organization unit") +
  labs(
    fill = "Organization unit",
    title = "Total Damage caused by Crime Incident") +
  theme(
    axis.ticks.x = element_blank(),
    axis.text.x = element_text(
      angle = -30,
      margin = margin(t = -5)
    ),
    axis.title.x = element_text(
      margin = margin(t = 15)
    ),
    axis.title.y = element_text(
      margin = margin(r = 15)
    ),
    plot.title = element_text(
      hjust = 0.5, vjust = 0.2, margin = margin(b = 20))
  )
```

```{r, echo=FALSE}
plot.dtotal
```

```{r}
plot.dpi <- ggplot(dat.total, aes(x = Základní.útvar.PČR..kde.došlo.k.činu,
                      y = total.damage.ratio)) +
  # geom_col(aes(fill = Základní.útvar.PČR..kde.došlo.k.činu)) +
  geom_col(aes(fill = total.incidents)) +
  geom_hline(  # total mean line
    aes(yintercept = mean(total.damage.ratio)),
    alpha = 0.4,
    linetype = 'dashed') +
  scale_fill_drsimonj(palette = 'cool', discrete = F) +
  ylab("Damage in CZK") +
  xlab("Brno organization unit") +
  labs(
    fill = "Organization unit",
    title = "Total Damage per Crime Incident") +
  theme(
    axis.ticks.x = element_blank(),
    axis.text.x = element_text(
      angle = -30,
      margin = margin(t = -5)
    ),
    axis.title.x = element_text(
      margin = margin(t = 15)
    ),
    axis.title.y = element_text(
      margin = margin(r = 15)
    ),
    plot.title = element_text(
      hjust = 0.5, vjust = 0.2, margin = margin(b = 20))
  )
```

```{r, echo=FALSE}
plot.dpi
```

Use cowplot to plot the partial plot

```{r}
plot_grid(plot.dtotal + theme(axis.text.x = element_blank(),
                              axis.title.x = element_blank()) + guides(fill = FALSE),
          plot.dpi + guides(fill = FALSE),
          align = 'hv', nrow = 2)
```


## Stolen goods

```{r}
dat.stolen.goods <- dat[dat$Předmět.zájmu.pachatele.nebo.vztah.pachatele.k.oběti...text != 'nic',]
```

There is quite a lot of levels

```{r}
## levels
levels(factor(dat.stolen.goods$Předmět.zájmu.pachatele.nebo.vztah.pachatele.k.oběti...text))
```

Filter robberies, thefts, financial frauds etc...

```{r}
# Refresher on our categories:

### x < 200 ~ "Violent crime",
### x < 300 ~ "Sexual harassment",
### x < 500 ~ "Robbery & Burglary",
### x < 580 ~ "Fraud",
### x < 630 ~ "Vandalism",
### x < 690 ~ "Drugs, Weapons & Dealing",
### x < 800 ~ "Other criminal activities",
### x < 900 ~ "Financial and economical criminal acts",
### x >= 900 ~ "Military crimes"

dat.stolen.goods %<>%
  filter(between(Takticko.statistická.klasifikace.činu, 300, 900))
```

Count number of crimes and the damage

```{r}
plot.stolen.goods <- ggplot(
  dat.stolen.goods, aes(x = factor(Předmět.zájmu.pachatele.nebo.vztah.pachatele.k.oběti...text))) +
  geom_bar(aes(fill = crime.category)) +
  theme(
    axis.line.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),

    axis.title.y = element_text(
      margin = margin(r = 10)
    ),
    axis.title.x = element_text(
      margin = margin(t = 10)
    ),

    panel.background = element_blank(),

    plot.title = element_text(hjust = 0.5)
  ) +
  xlab("Stolen goods") +
  ylab("Number of crimes") +
  labs(
    fill = "Crime category",
    title = "Crime in Brno by Stolen Goods"
  )
```

Plot interactively (if applicable)

> ⚠ NOTE: will be rendered as JPEG in embedded markdown notebooks) ⚠

```{r}
plot.stolen.goods  # (ggplotly(plot.stolen.goods))
```

## And what interests us the most.. which part of Brno has the highest criminality in general?

```{r}
plot.itotal <- ggplot(dat, aes(x = factor(str_remove(Základní.útvar.PČR..kde.došlo.k.činu, "OOP BRNO -")))) +
  geom_bar(aes(fill = crime.category)) +
  theme(
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),

    axis.text.x = element_text(
      angle = -30,
      margin = margin(t = -5, b = 20)
    ),

    panel.background = element_blank(),

    plot.title = element_text(hjust = 0.5)
  ) +
  ylab("Number of crimes") +
  xlab("Brno OOP unit") +
  labs(
    fill = "Crime category",
    title = "Crime in Brno by Category"
  )
```

Much better to explore these interactively (if possible)

```{r, echo=FALSE}
plot.itotal  # ggplotly(plot.itotal)
```

### Geospatial

```{r}
library(sf)

shapefile  <- st_read("data/shape_files/moc/Městské_obvody_a_městské_části__polygony.shp")
shape.brno <- shapefile %>%
  filter(str_detect(NAZ_ZUJ, 'Brno')) %>%
  fortify()

### transform and unify municipality names
shape.brno$NAZ_ZUJ %<>%
  sapply(str_remove, 'Brno-') %>%
  sapply(str_to_title) %>%
  str_replace("^.*?-", "")
```

Now the tricky part. OOP are special police departments and there is not any shape file for them.
We can try to merge them based on the municipality units.

```{r}
df.brno.oop <- readxl::read_xls("data/brno_data/Seznam_ulic_mesta_Brna_s_odkazy.xls")
df.brno.oop %<>%
  mutate(OOP = str_remove(OOP, pattern="OOP ")) %>%
  mutate(OOP = str_to_title(OOP)) %>%
  select(OOP, NAZ_ZUJ = `Městská část`) %>%
  mutate(NAZ_ZUJ = str_to_title(NAZ_ZUJ)) %>%
  mutate(NAZ_ZUJ = str_replace(NAZ_ZUJ, "^.*?-", "")) %>%
  select(OOP, NAZ_ZUJ) %>%
  unique()


df.brno <- merge(shape.brno, df.brno.oop, by = "NAZ_ZUJ")

# pre-process
dat.total$Základní.útvar.PČR..kde.došlo.k.činu %<>%
  sapply(str_remove, "OOP ") %>%
  sapply(str_replace, " - ", "-") %>%
  sapply(str_to_title)

dat.total %<>%
  rename(OOP = Základní.útvar.PČR..kde.došlo.k.činu)

# merge with brno moc
df.brno %<>% merge(dat.total, by = "OOP") %>%
  transform(OOP = factor(OOP)) %>%
  unique()
```

```{r}
ggplot(data = df.brno) +
  geom_sf(aes(fill = NAZ_ZUJ, geometry = geometry)) +
  coord_sf(datum = NA) +
  # scale_fill_drsimonj(discrete = T) +
  labs(
    title = "Brno municipality units",
    fill = 'Municipality unit') +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))
```

Fill by total damage

```{r}
df.brno %<>%
  group_by(NAZ_ZUJ) %>%
  mutate(incidents.in.ZUJ = sum(total.incidents)) %>%
  mutate(damage.in.ZUJ = sum(total.damage))
```

```{r}
plot.geom.damage <- ggplot(data = df.brno, aes(label = NAZ_ZUJ)) +
  geom_sf(aes(fill = damage.in.ZUJ, geometry = geometry)) +
  geom_label(
    aes(x = SX, y = SY, size=1, alpha=0.4), label.size = 0.10, show.legend = F) +
  coord_sf(datum = NA) +
  scale_fill_drsimonj(discrete = F) +
  labs(
    title = "Crime in Brno by damage caused",
    fill = 'Damage in CZK (Millions)') +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r, echo=FALSE}
plot.geom.damage
```

Fill by total number of crimes

```{r}
plot.geom.incidents <- ggplot(data = df.brno, aes(label = NAZ_ZUJ)) +
  geom_sf(aes(fill = incidents.in.ZUJ, geometry = geometry)) +
  geom_label(
    aes(x = SX, y = SY, size=1, alpha=0.4), label.size = 0.10, show.legend = F) +
  coord_sf(datum = NA) +
  scale_fill_drsimonj(discrete = F) +
  labs(
    title = "Crime in Brno",
    fill = 'Number of incidents') +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```

```{r, echo=FALSE}
plot.geom.incidents
```

## And the last, my favourite -- Wordcloud

```{r}
library(ggwordcloud)
```

Prepare the data

```{r}
dat.stolen.goods %<>%
  group_by(Předmět.zájmu.pachatele.nebo.vztah.pachatele.k.oběti...text) %>%
  mutate(total.crimes = n(), total.damage = sum(Celková.způsobená.škoda.v.Kč)) %>%
  select(
    items = Předmět.zájmu.pachatele.nebo.vztah.pachatele.k.oběti...text,
    total.crimes,
    total.damage
  ) %>%
  transform(items = factor(items)) %>%
  unique()

# truncate words
dat.stolen.goods %<>%
  mutate(items.trunc = sapply(as.character(items), str_trunc, width = 10))

# get the quantile
q.90 <- dat.stolen.goods %>%
  pull(total.crimes) %>%
  quantile(.90)

# set the angle randomly + based on the quantile
dat.stolen.goods %<>%
  mutate(angle = sample(seq(-60, 60, 15), n(), replace = T))
dat.stolen.goods[which(dat.stolen.goods$total.crimes > q.90),]$angle <- 0
```

Plot the wordcloud

```{r}
set.seed(42)
ggplot(dat.stolen.goods, aes(label = items.trunc,
                             size  = total.crimes,
                             color = items.trunc)) +
  geom_text_wordcloud_area() +
  scale_size_area(max_size = 30) +
  theme_minimal()
```

Color by total damage (DevConf style)

```{r}
set.seed(42)
ggplot(dat.stolen.goods, aes(label = items.trunc,
                             size  = total.crimes,
                             color = total.damage)) +
  geom_text_wordcloud_area() +
  scale_size_area(max_size = 30) +
  theme_minimal() +
  scale_color_gradient(low = '#8e82e4', high = 'purple')
```


Lets get fancy

```{r}
plot.wordcloud <- ggplot(dat.stolen.goods, aes(label = items.trunc,
                                               size  = total.crimes,
                                               color = total.damage,
                                               angle = angle)) +
  geom_text_wordcloud_area(
    mask = png::readPNG("img/brno_znak.filled.png"),
    rm_outside = T  # some of those will get removed, no time to figgle parameters
  ) +
  scale_size_area(max_size = 24) +
  theme_minimal() +
  scale_color_gradient(low = 'darkred', high = 'red')
```

```{r, echo=FALSE}
plot.wordcloud
```

```{r}
brno.title <- ggdraw() + draw_label(
  "BRNO",
  colour = 'red',
  size = 60,
  fontfamily = 'Overpass',
  fontface = 'bold'
)

brno.emblem <- ggdraw() +
  draw_image(png::readPNG("img/brno_znak.red.transparent.png")) +
  draw_plot(plot.wordcloud)
```

```{r}
plot_grid(brno.title, brno.emblem, ncol = 1, rel_heights = c(0.2, 1))
```



Save the processed data (if applicable)

```{r}
# write_csv(df.brno, 'brno.crime.incidents.csv')
```
